Penalized linear mixed models for structured genetic data

### 2021-03-15 main.tex
adding renamed figures per journal guidelines
Adding ESL ref

\begin{equation}
    Q(\bbeta | \lam) = \frac{1}{2n}|| \by -  \beta_0 - \bX \bbeta ||_2^2 + \lam ||\bbeta||_1,
\end{equation}
where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm. The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data (here, assuming that observations are independent). The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values. The tuning parameter, $\lambda$, controls the tradeoff between these two components. For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0. At the other extreme, if $\lam = 0$ the lasso solution coincides with that of classical ordinary least squares regression. The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).

@@

\pb{additional references?  Elements of statistical learning?}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{figures/beta_mse.png}
    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations. Mean and standard error (SE) bars displayed for 1000 replications.}
    \label{fig:mse}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}
\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
\label{fig:pop_inf}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
     \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
    \label{fig:eta_beta_mse}
\end{figure}

@@

Selection accuracy was evaluated using selection precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
\caption{\pb{I feel like we should stop this at 200.  The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_mse_p1=100.png}
\caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_mse}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr_p1=100.png}
\caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs. The plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular selection number, with shaded region indicating SE bounds. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_pc1_corr}
\end{figure}

>>>>

\begin{equation}
    Q(\bbeta | \lam) = \frac{1}{2n}|| \by -  \beta_0 - \bX \bbeta ||_2^2 + \lam ||\bbeta||_1,
\end{equation}
where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm. The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data (here, assuming that observations are independent). The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values. The tuning parameter, $\lambda$, controls the tradeoff between these two components. For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0. At the other extreme, if $\lam = 0$ the lasso solution coincides with that of classical ordinary least squares regression. The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox and logistic regression \citep{tibshirani1997lasso, friedman2001elements, shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).

@@

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{figures/figure_01.png}
    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations. Mean and standard error (SE) bars displayed for 1000 replications.}
    \label{fig:mse}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_02.png}
\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
\label{fig:pop_inf}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/figure_03.png}
    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{figures/figure_04.png}
     \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
    \label{fig:eta_beta_mse}
\end{figure}

@@

Selection accuracy was evaluated using selection precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_05.png}
\caption{\pb{I feel like we should stop this at 200.  The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_06.png}
\caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_mse}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_07.png}
\caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs. The plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular selection number, with shaded region indicating SE bounds. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_pc1_corr}
\end{figure}
